---
title: "Introduction to the covid19italy Datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The covid19italy R package provides a tidy format dataset of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) pandemic outbreak in Italy. The package includes the following three datasets:

- `italy_total` - daily summary of the outbreak on the national level
- `italy_region` - daily summary of the outbreak on the region level
- `italy_province` - daily summary of the outbreak on the province level

The data was pull from [Italy Department of Civil Protection](http://www.protezionecivile.it/)

## Installation

You can install the released version of covid19italy from [CRAN](https://cran.r-project.org/package=covid19italy) with:

``` r
install.packages("covid19italy")
```

Or, install the most recent version from [GitHub](https://github.com/RamiKrispin/covid19italy) with:

``` r
# install.packages("devtools")
devtools::install_github("RamiKrispin/covid19Italy")
```
## Data refresh


The **covid19italy** package dev version is been updated on a daily bases. The `update_data` function enables a simple refresh of the installed package datasets with the most updated version on Github:

``` r
library(covid19italy)

update_data()
```

Note: must restart the R session to have the updates available

## Italy summary

The `italy_total` dataset provides an overall summary of the cases in Italy since the beginning of the covid19 outbreak since February 24, 2020. The dataset contains the following fields:

* `date` - timestamp, a `Date` object
* `hospitalized_with_symptoms` - daily number of patients hospitalized with symptoms
* `intensive_care` - daily number of patients on intensive care
* `total_hospitalized` - daily total number of patients hospitalized (`hospitalized_with_symptoms` + `intensive_care`)
* `home_confinement` - daily number of people under home confinement
* `cumulative_positive_cases` - a daily snapshot of the number of positive cases
* `daily_positive_cases` - daily new positive cases
* `daily_cases` - daily new positive, recovered, and death cases
* `recovered` - total number of recovered cases (cumulative)
* `death` - total number of death cases (cumulative)
* `positive_clinical_activity` - positive cases emerged from clinical activity
* `positive_surveys_tests` - positive cases emerging from surveys and tests, planned at national or regional level
* `cumulative_cases` - total number of positive cases (cumulative)
* `total_tests` - total number of tests performed (cumulative)


```{r}
library(covid19italy)

data(italy_total)

str(italy_total)

head(italy_total)
```



## Italy region level

The `italy_region` dataset provides an overall summary of the cases in Italy's regions. The dataset contains the following fields:

* `date` - timestamp, a `Date` object
* `region_code` - the region code
* `region_name` - the region name
* `lat` - region latitude coordinate
* `long` - region longitude coordinate
* `hospitalized_with_symptoms` - daily number of patients hospitalized with symptoms
* `intensive_care` - daily number of patients on intensive care
* `total_hospitalized` - daily total number of patients hospitalized (`hospitalized_with_symptoms` + `intensive_care`)
* `home_confinement` - daily number of people under home confinement
* `cumulative_positive_cases` - a daily snapshot of the number of positive cases
* `daily_positive_cases` - daily new positive cases
* `daily_cases` - daily new positive, recovered, and death cases
* `recovered` - total number of recovered cases (cumulative)
* `death` - total number of death cases (cumulative)
* `positive_clinical_activity` - positive cases emerged from clinical activity
* `positive_surveys_tests` - positive cases emerging from surveys and tests, planned at national or regional level
* `cumulative_cases` - total number of positive cases, recovered, and death (cumulative)
* `total_tests` - total number of tests performed (cumulative)
* `region_spatial` - the spatial region names as in the output of the `ne_states` function from the **rnaturalearth** package  


```{r}
data(italy_region)

str(italy_region)

head(italy_region)
```

## Italy province level

The `italy_region` dataset provides an overall summary of the cases in Italy's regions. The dataset contains the following fields:

* `date` - timestamp, a `Date` object
* `region_code` - the region code
* `region_name` - the region name
* `province_code` - the province code
* `province_name` - the province name
* `province_abb` - the province abbreviation
* `lat` - province latitude coordinate
* `long` - province longitude coordinate
* `total_cases` - total number of positive cases (cumulative)
* `new_tests` - daily number of positive cases
* `province_spatial` - the spatial province names as in the output of the `ne_states` function from the **rnaturalearth** package


```{r}
data(italy_province)

str(italy_province)

head(italy_province)
```
